Select-based random access \\ to variable-byte encodings

### 2020-04-17 gradu/gradu.tex
mods
new stuff
Updated pdf
more text

\usepackage{lmodern}         % Font package, again in some systems.
\usepackage{textcomp}        % Package for special symbols
\usepackage[pdftex]{color, graphicx} % For pdf output and jpg/png graphics
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage[pdftex, plainpages=false]{hyperref} % For hyperlinks and pdf metadata
\usepackage{fancyhdr}        % For nicer page headers
\usepackage{tikz}            % For making vector graphics (hard to learn but powerful)
%\usepackage{wrapfig}        % For nice text-wrapping figures (use at own discretion)
\usepackage{amsmath, amssymb} % For better math
\usepackage{algorithm}  
\usepackage[noend]{algpseudocode}
\singlespacing               %line spacing options; normally use single

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% STEP 2:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Set up personal information for the title page and the abstract form.
%% Replace parameters with your information.
\title{something about variable-byte encoding}

@@

\classification{\protect{\ \\
\  General and reference $\rightarrow$ Document types  $\rightarrow$ Surveys and overviews\  \\
\  Applied computing  $\rightarrow$ Document management and text processing  $\rightarrow$ Document management $\rightarrow$ Text editing
}}

@@

\begin{abstract}
Use this otherlanguage environment to write your abstract in another language if needed.

Topics are classified according to the ACM Computing Classification System
(CCS), see 
\url{https://www.acm.org/about-acm/class}:
check command \verb+\classification{}+. A small set of paths (1--3) should be used, starting from any top nodes
referred to bu the root term CCS leading to the leaf nodes. The elements
in the path are separated by right arrow, and emphasis of each element individually can be indicated
by the use of bold face for high importance or italics for intermediate
level. The combination of individual boldface terms may give the reader
additional insight. 
\end{abstract}

>>>>

\usepackage{lmodern}         % Font package, again in some systems.
\usepackage{textcomp}        % Package for special symbols
\usepackage[pdftex]{color, graphicx} % For pdf output and jpg/png graphics
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage[pdftex, plainpages=false]{hyperref} % For hyperlinks and pdf metadata
\usepackage{fancyhdr}        % For nicer page headers
\usepackage{tikz}             % For making vector graphics (hard to learn but powerful)
\usetikzlibrary{arrows.meta,scopes,chains,matrix}
%\usepackage{wrapfig}        % For nice text-wrapping figures (use at own discretion)
\usepackage{amsmath, amssymb} % For better math
\usepackage{algorithm}  
\usepackage[noend]{algpseudocode}
\usepackage{pgfplots} 
\singlespacing               %line spacing options; normally use single

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% STEP 2:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Set up personal information for the title page and the abstract form.
%% Replace parameters with your information.
\title{Select-based random access \\ to variable-byte encodings}

@@

\classification{\protect{\ \\
\  Information systems $\rightarrow$ Data management systems  $\rightarrow$ Data structures $\rightarrow$ Data layout $\rightarrow$ Data compression \  \\
\  probably $\rightarrow$ something $\rightarrow$ else $\rightarrow$ too \\
}}

@@

\begin{abstract}

Enormous datasets are a common occurence today and compressing them is often beneficial. Variable-byte encoding is a method for compressing integers of different byte lenghts. It removes the zero bytes 
and adds an additional \textit{continuation bit} to each byte to denote whether the compressed integer continues to the next byte or not. Two bit array data structures, $rank$ and $select$  are built on top of the array of 
continuation bits. These data structures enable a fast direct access to the elements in the compressed data. The $rank$ implementation performs better on data containing mostly small integers. The $select$ 
implementation benefits on larger integers. An experiment on fetching a subarray is explained and implemented. The $select$ implementation has significant advantages on subarray fetching due to how the data 
is compressed.

\end{abstract}

### 2020-04-17 gradu/msc_english_contents.tex
more text

Variable-byte encoding is a method for compressing integers of different byte length. An additional $continuous bit$ is added to the data to denote whether the compressed integer
is continued to the next byte or not. A light-weight data structure is built over the continuous bit array. The fact that the compressed data blocks are of the same size is used
with the data structure to allow direct access to the compressed variable-byte data.

@@

Elias Delta and Gamma codes \citep{Eli75} are popular encoding methods for the aforementioned data sets. Their encoding process assigns short bit sequences to small integers, 
which is why they outperform in trems of compression VB encoding on datasets containing a lot of small integers \citep{Wil99}, such as inverted indexes \citep{Anh05, Pib19}. In this thesis, 
the data sets used for experimentation will have the focus on slightly larger, unordered numbers.

@@

\begin{figure}[ht]
\centering
\begin{algorithmic}
\State $i \gets $ wanted index
\State $A \gets $ block byte array
\State $B \gets $ continuation bit array
\State $begin \gets \Call{Select}{index}$
\State $len \gets \Call{CalculateLength}{begin}$
\State $mask \gets \Call{ByteMask}{len}$
\State $word \gets A[begin]$ \Comment{read a word from the byte array}
\State $word \gets word \mathbin{\&} mask$

@@

>>>>

Variable-byte encoding is a method for compressing integers of different byte length. An additional \textit{continuous bit} is added to the data to denote whether the compressed integer
is continued to the next byte or not. A light-weight data structure is built over the continuous bit array. The fact that the compressed data blocks are of the same size is used
with the data structure to allow direct access to the compressed variable-byte data.

@@

Elias Delta and Gamma codes \citep{Eli75} are popular encoding methods for the aforementioned data sets. Their encoding process assigns short bit sequences to small integers, 
which is why they outperform in terms of compression VB encoding on datasets containing a lot of small integers \citep{Wil99}, such as inverted indexes \citep{Anh05, Pib19}. In this thesis, 
the data sets used for experimentation will have the focus on slightly larger, unordered numbers.

@@

\begin{figure}[ht]
\centering
\begin{algorithmic}
\State $i \gets $ wanted index
\State $A \gets $ block byte array
\State $B \gets $ continuation bit array
\State $begin \gets \Call{Select}{B, index}$
\State $len \gets \Call{CalculateLength}{begin}$
\State $mask \gets \Call{ByteMask}{len}$
\State $word \gets A[begin]$ \Comment{read a word from the byte array}
\State $word \gets word \mathbin{\&} mask$

@@

 - remove lists from algorithms and data sets in results, explain whe data sets were roughly chosen like that
